### 2016-09-01 R/ITHIM.Rmd
eval = TRUE for whole script

```{r parsely, eval = FALSE}
convertToHHS <- function(states){
  HHSRegions <- matrix(c("CT","ME","MA","NH","RI","VT","NJ","NY","DE","DC","MD","PA","VA","WV","AL","FL","GA","KY","MS","NC","SC","TN","IL","IN","MI","MN","OH","WI","AR","LA","NM","OK","TX","IA","KS","MO","NE","CO","MT","ND","SD","UT","WY","AZ","CA","HI","NV","AK","ID","OR","WA",1,1,1,1,1,1,2,2,3,3,3,3,3,3,4,4,4,4,4,4,4,4,5,5,5,5,5,5,6,6,6,6,6,7,7,7,7,8,8,8,8,8,8,9,9,9,9,10,10,10,10,"HHS1","HHS1","HHS1","HHS1","HHS1","HHS1","HHS2","HHS2","HHS3","HHS3","HHS3","HHS3","HHS3","HHS3","HHS4","HHS4","HHS4","HHS4","HHS4","HHS4","HHS4","HHS4","HHS5","HHS5","HHS5","HHS5","HHS5","HHS5","HHS6","HHS6","HHS6","HHS6","HHS6","HHS7","HHS7","HHS7","HHS7","HHS8","HHS8","HHS8","HHS8","HHS8","HHS8","HHS9","HHS9","HHS9","HHS9","HHS10","HHS10","HHS10","HHS10"), nrow = 51, ncol = 3)
  StateToHHS <- HHSRegions[,3]
  names(StateToHHS) <- HHSRegions[,1]
  return(StateToHHS[states])
}

@@

```{r NHTS, eval = FALSE}
DAYV2PUB <- read.csv(file = "~/GHI/data/NHTS/Ascii/DAYV2PUB.CSV", na.strings = c("-9","-8", "-7", "-1"))
names(DAYV2PUB) <- c("HOUSEID","PERSONID","FRSTHM","OUTOFTWN","ONTD_P1","ONTD_P2","ONTD_P3","ONTD_P4","ONTD_P5","ONTD_P6","ONTD_P7","ONTD_P8","ONTD_P9","ONTD_P10","ONTD_P11","ONTD_P12","ONTD_P13","ONTD_P14","ONTD_P15","TDCASEID","HH_HISP","HH_RACE","DRIVER","R_SEX","WORKER","DRVRCNT","HHFAMINC","HHSIZE","HHVEHCNT","NUMADLT","FLAG100","LIF_CYC","TRIPPURP","AWAYHOME","CDIVMSAR","CENSUS_D","CENSUS_R","DROP_PRK","DRVR_FLG","EDUC","ENDTIME","HH_ONTD","HHMEMDRV","HHRESP","HHSTATE","HHSTFIPS","INTSTATE","MSACAT","MSASIZE","NONHHCNT","NUMONTRP","PAYTOLL","PRMACT","PROXY","PSGR_FLG","R_AGE","RAIL","STRTTIME","TRACC1","TRACC2","TRACC3","TRACC4","TRACC5","TRACCTM","TRAVDAY","TREGR1","TREGR2","TREGR3","TREGR4","TREGR5","TREGRTM","TRPACCMP","TRPHHACC","TRPHHVEH","TRPTRANS","TRVL_MIN","TRVLCMIN","TRWAITTM","URBAN","URBANSIZE","URBRUR","USEINTST","USEPUBTR","VEHID","WHODROVE","WHYFROM","WHYTO","WHYTRP1S","WRKCOUNT","DWELTIME","WHYTRP90","TDTRPNUM","TDWKND","TDAYDATE","TRPMILES","WTTRDFIN","VMT_MILE","PUBTRANS","HOMEOWN","HOMETYPE","HBHUR","HTRESDN","HTHTNRNT","HTPPOPDN","HTEEMPDN","HBRESDN","HBHTNRNT","HBPPOPDN","GASPRICE","VEHTYPE","HH_CBSA","HHC_MSA")

@@

```{r loadData, eval = FALSE}
NHTS.person <- readRDS(file = "./data/NHTS.person.rds")
NHTS.trips <- readRDS(file = "./data/NHTS.trips.rds")
```

@@

```{r summary, fig.width = 20, fig.height = 10, eval = TRUE}
summary(with(subset(NHTS.trips, AT == "walk"), TRPMILES))
summary(with(subset(NHTS.trips, AT == "cycle"), TRPMILES))
summary(with(subset(NHTS.trips, AT == "other"), TRPMILES))
summary(with(subset(NHTS.trips, AT == "walk"), TRVLCMIN))
summary(with(subset(NHTS.trips, AT == "cycle"), TRVLCMIN))
summary(with(subset(NHTS.trips, AT == "other"), TRVLCMIN))

@@

```{r subset, eval = TRUE, fig.width = 10}
NHTS.person <- subset( NHTS.person, MSA == TRUE )
NHTS <- merge(NHTS.person, NHTS.trips, by = c("HOUSEID","PERSONID"))
barplot(with(NHTS,table(HHS)))
```

@@

```{r split, eval = FALSE}
listOfLists <- lapply(split(NHTS, NHTS$HHS),function(x) lapply(split(x, x$SEX), function(x) split(x, x$AGE)))
```

@@

```{r count, eval = FALSE}
counts <- lapply(listOfLists, function(x) lapply(x, function(x) lapply(x, countPeople)))
walkLength <- lapply(listOfLists, function(x) lapply(x, function(x) lapply( x, tripTime, mode = "walk")))
cycleLength <- lapply(listOfLists, function(x) lapply(x, function(x) lapply( x, tripTime, mode = "cycle")))
countMatrix <- lapply(counts, reformatData)
walkMatrix <- lapply(walkLength, reformatData)
cycleMatrix <- lapply(cycleLength, reformatData)
meanWalkList <- mapply(walkMatrix,   countMatrix, FUN = function(x,y) x/y*7, SIMPLIFY = FALSE)
meanCycleList <- mapply(cycleMatrix, countMatrix, FUN = function(x,y) x/y*7, SIMPLIFY = FALSE)
RwtList <- lapply( meanWalkList, function(x) ifelse(x == 0, 0.01, x) )
RctList <- lapply( meanCycleList, function(x) ifelse(x == 0, 0.01, x) )
RwtList <- lapply( RwtList, function(x) x/x[3,2] )
RctList <- lapply( RctList, function(x) x/x[3,2] )
muwtList <- lapply( meanWalkList, function(x) x[3,2] )
muctList <- lapply( meanCycleList, function(x) x[3,2] )
```

@@

```{r CDC, eval = FALSE}
wonder <- read.csv("./../data/CDCwonder/HHSmetro_Age_sex_2010-14.csv", header=T)
names(wonder) <- c("HHSRegion", "HHSRegionCode", "gender", "genderCode","WONDERage","WONDERageCode","ICD10",  "ICD10Code", "deaths5yr", "population5yr", "crudeRate")

@@

```{r ITHIMCause, eval = FALSE}
ITHIMcauseKey <- c("ColonCancer","LungCancer","BreastCancer","Diabetes","Dementia","IHD","InflammatoryHD","InflammatoryHD","HHD","Stroke","AcuteRespInfections","AcuteRespInfections","RespiratoryDisease","RespiratoryDisease","RTIs","RTIs")
names(ITHIMcauseKey) <- unique(wonder$ICD10)
wonder$ITHIMcause <- ITHIMcauseKey[as.character(wonder$ICD10)]
```
```{r recodeAge, eval = FALSE}
ITHIMageKey <- c("ageClass1", "ageClass1", "ageClass2", "ageClass2", "ageClass3", "ageClass3", "ageClass3", "ageClass4", "ageClass4", "ageClass4", "ageClass5", "ageClass5", "ageClass5", "ageClass6", "ageClass6", "ageClass7", "ageClass7", "ageClass8", "ageClass8","ageClass8", "ageClass8", "ageClass8")
names(ITHIMageKey) <- unique(wonder$WONDERage)
wonder$ITHIMage <- ITHIMageKey[as.character(wonder$WONDERage)]
```

@@

```{r wonder, eval = FALSE}
wonder <- wonder[, c("HHSRegionCode","gender","ITHIMage","ITHIMcause","deaths5yr","population5yr")]
depress <- expand.grid(HHSRegionCode=unique(wonder$HHSRegionCode), gender=unique(wonder$gender),ITHIMage = unique(wonder$ITHIMage), ITHIMcause = "Depression", deaths5yr = "0", population5yr = "0", stringsAsFactors=T)
wonder <- rbind(wonder, depress)
```

@@

```{r suppressed, eval = FALSE}
wonder <- within( wonder, {
    newDeaths <- as.numeric(ifelse(deaths5yr == "Suppressed", sample(0:9), as.numeric(as.character(deaths5yr))))
    })
```

@@

```{r summary2, eval = FALSE}
wonderSummary <- ddply(wonder, .(HHSRegionCode, gender, ITHIMage, ITHIMcause), summarise,
                       numberOfDeaths5yr = sum(as.numeric(as.character(newDeaths)), na.rm=T)
)
```

@@

```{r singleyear, eval = FALSE}
pop14 <- read.csv("./../data/CDCwonder/HHSmetro_Age_sex_2014_AllCause.csv", header=T)

@@

```{r sloppy, eval = FALSE}
ITHIMageKey2 <- c("ageClass8", "ageClass1", "ageClass7", "ageClass7", "ageClass6", "ageClass6", "ageClass1", "ageClass5", "ageClass5","ageClass5","ageClass4", "ageClass3", "ageClass4", "ageClass2", "ageClass2", "ageClass3", "ageClass4", "ageClass3","ageClass8")
names(ITHIMageKey2) <- unique(pop14$WONDERage)
pop14$ITHIMage <- ITHIMageKey2[as.character(pop14$WONDERage)]
```

```{r foop, eval = FALSE}
pop14 <-  pop14[, c("HHSRegionCode","gender","ITHIMage","deaths2014AllCause","population2014")]
pop14$deaths2014 <- as.numeric(ifelse(pop14$deaths2014 == "Suppressed", sample(0:9), as.numeric(as.character(pop14$deaths2014))))

@@

```{r USGBD, eval = FALSE}
USGBD <- read.csv("./../data/CDCwonder/USGBD.csv", header=T)
USGBD$ITHIMage <- as.character(USGBD$ITHIMage)
USGBD$ITHIMcause <- as.character(USGBD$ITHIMcause)

@@

```{r burdenList, eval = FALSE}
GBDList <- lapply(split(burdenOut, burdenOut$HHS),function(x) lapply(split(x, x$disease), function(x) split(x, x$sex)))
#saveRDS(GBDList, file = "../data/GBD.rds")
```

@@

```{r newMETs, echo = TRUE, eval = FALSE}
newMETsFile <- "~/GHI/data/ATUS/ATUS_to_METS_key.csv"
newMETs <- read.csv(file = newMETsFile, stringsAsFactors = FALSE, colClasses = c("character","character","character","character","numeric","character","character","character"))
dataColumns <- c(1,5)
D <- newMETs[,dataColumns]
metaData <- newMETs[,-dataColumns]
newMETs <- list(Data = D, metaData = metaData)
activityCode <- ifelse(nchar(newMETs$Data$X6.digit.activity.code)==5,paste0("t0",newMETs$Data$X6.digit.activity.code),paste0("t",newMETs$Data$X6.digit.activity.code))

@@

```{r readnewMETs, echo = TRUE, eval = FALSE}
cpsFile <- "~/GHI/data/ATUS/atuscps_0311.dat"
cps <- read.table(cpsFile, header=T, skip=0, sep="," )
cps <- within(cps, {
    TUCASEID <- as.character(TUCASEID)
    TULINENO <- as.character(TULINENO)
    GEREG <- as.factor(GEREG)
    })
cps <- subset(cps, TULINENO == "1")

@@

```{r metKey, echo = TRUE, eval = FALSE}
allActivities <- colnames(summaryList$Data)
nonTravelActivities <- rownames(newMETs$metaData)[which(newMETs$metaData$Travel.related..Y.N. == "N")]
nonTravelActivities <- nonTravelActivities[which(nonTravelActivities %in% allActivities)]

@@

```{r parameters, eval = FALSE}
(parameters <- ddply(D, .(SEX, AGECAT), summarise,
             mean = mean(METs, na.rm=T),
             sd = sd(METs, na.rm=T)
             ))

@@

```{r foobar, eval = TRUE}
sumDALY <- function(ITHIM.baseline, ITHIM.scenario){
    return(sum(unlist(compareModels(ITHIM.baseline, ITHIM.scenario)$daly.delta)))
    }
```

```{r ITHIMObjs, eval = TRUE, fig.width = 20, fig.height = 10}
compute <- FALSE

@@

```{r meanMatrices, eval = TRUE, fig.width = 20, fig.height = 10, warning = FALSE, message = FALSE}

@@

```{r CRA, eval = TRUE}

@@

```{r instantiate, eval = FALSE}
ITHIM.baseline <- createITHIM(vision = "baseline", region = "SFBayArea")
ITHIM.scenario <- createITHIM(vision = "scenario", region = "SFBayArea")
```

@@

```{r RRs, eval = FALSE}
comparativeRisk <- compareModels(ITHIM.baseline,ITHIM.scenario)
```

### Deaths
```{r show, echo = FALSE, eval = FALSE}
matrix(unlist(lapply(comparativeRisk$dproj.delta,function(x) lapply(x, function(x) sum(x[-(1:2)]) ))),nrow=5,ncol=2,byrow=TRUE,dimnames=list(names(comparativeRisk$dproj.delta),c("M","F")))
```
### DALY
```{r show2, echo = FALSE, eval = FALSE}
matrix(unlist(lapply(comparativeRisk$daly.delta,function(x) lapply(x, function(x) sum(x[-(1:2)]) ))),nrow=5,ncol=2,byrow=TRUE,dimnames=list(names(comparativeRisk$dproj.delta),c("M","F")))
```
### YLD
```{r show3, echo = FALSE, eval = FALSE}
matrix(unlist(lapply(comparativeRisk$yld.delta,function(x) lapply(x, function(x) sum(x[-(1:2)]) ))),nrow=5,ncol=2,byrow=TRUE,dimnames=list(names(comparativeRisk$dproj.delta),c("M","F")))
```
### YLL
```{r show4, echo = FALSE, eval = FALSE}
matrix(unlist(lapply(comparativeRisk$yll.delta,function(x) lapply(x, function(x) sum(x[-(1:2)]) ))),nrow=5,ncol=2,byrow=TRUE,dimnames=list(names(comparativeRisk$dproj.delta),c("M","F")))
```

@@

```{r instantiate2, eval = FALSE}
ITHIM.baseline <- createITHIM(vision = "baseline", region = "national")
ITHIM.scenario <- createITHIM(vision = "scenario", region = "national")
```

```{r RRs2, eval = FALSE}
comparativeRisk <- compareModels(ITHIM.baseline,ITHIM.scenario)
```

### Deaths
```{r show222, echo = FALSE, eval = FALSE}
matrix(unlist(lapply(comparativeRisk$dproj.delta,function(x) lapply(x, function(x) sum(x[-(1:2)]) ))),nrow=5,ncol=2,byrow=TRUE,dimnames=list(names(comparativeRisk$dproj.delta),c("M","F")))
```
### DALY
```{r show22, echo = FALSE, eval = FALSE}
matrix(unlist(lapply(comparativeRisk$daly.delta,function(x) lapply(x, function(x) sum(x[-(1:2)]) ))),nrow=5,ncol=2,byrow=TRUE,dimnames=list(names(comparativeRisk$dproj.delta),c("M","F")))
```
### YLD
```{r show23, echo = FALSE, eval = FALSE}
matrix(unlist(lapply(comparativeRisk$yld.delta,function(x) lapply(x, function(x) sum(x[-(1:2)]) ))),nrow=5,ncol=2,byrow=TRUE,dimnames=list(names(comparativeRisk$dproj.delta),c("M","F")))
```
#### YLL
```{r show24, echo = FALSE, eval = FALSE}
matrix(unlist(lapply(comparativeRisk$yll.delta,function(x) lapply(x, function(x) sum(x[-(1:2)]) ))),nrow=5,ncol=2,byrow=TRUE,dimnames=list(names(comparativeRisk$dproj.delta),c("M","F")))
```

@@

```{r example, eval = FALSE}
example("ITHIM-package")
```

@@

```{r vision, eval = FALSE}
ITHIM.vision <- updateITHIM(ITHIM.scenario, "muwt", 2*ITHIM.scenario$parameters$muwt) # Double the mean walking time
comparativeRisk <- compareModels(ITHIM.baseline,ITHIM.vision)
```
#### YLL
```{r show233, echo = FALSE, eval = FALSE}
matrix(unlist(lapply(comparativeRisk$yll.delta,function(x) lapply(x, function(x) sum(x[-(1:2)]) ))),nrow=5,ncol=2,byrow=TRUE,dimnames=list(names(comparativeRisk$dproj.delta),c("M","F")))
```

@@

```{r params, eval = FALSE}
names(ITHIM.baseline$parameters)
ITHIM.baseline$parameters$NonTravelMETs
ITHIM.baseline$parameters$Rwt
```

>>>>

```{r parsely, eval = TRUE}
convertToHHS <- function(states){
  HHSRegions <- matrix(c("CT","ME","MA","NH","RI","VT","NJ","NY","DE","DC","MD","PA","VA","WV","AL","FL","GA","KY","MS","NC","SC","TN","IL","IN","MI","MN","OH","WI","AR","LA","NM","OK","TX","IA","KS","MO","NE","CO","MT","ND","SD","UT","WY","AZ","CA","HI","NV","AK","ID","OR","WA",1,1,1,1,1,1,2,2,3,3,3,3,3,3,4,4,4,4,4,4,4,4,5,5,5,5,5,5,6,6,6,6,6,7,7,7,7,8,8,8,8,8,8,9,9,9,9,10,10,10,10,"HHS1","HHS1","HHS1","HHS1","HHS1","HHS1","HHS2","HHS2","HHS3","HHS3","HHS3","HHS3","HHS3","HHS3","HHS4","HHS4","HHS4","HHS4","HHS4","HHS4","HHS4","HHS4","HHS5","HHS5","HHS5","HHS5","HHS5","HHS5","HHS6","HHS6","HHS6","HHS6","HHS6","HHS7","HHS7","HHS7","HHS7","HHS8","HHS8","HHS8","HHS8","HHS8","HHS8","HHS9","HHS9","HHS9","HHS9","HHS10","HHS10","HHS10","HHS10"), nrow = 51, ncol = 3)
  StateToHHS <- HHSRegions[,3]
  names(StateToHHS) <- HHSRegions[,1]
  return(StateToHHS[states])
}

@@

```{r NHTS, eval = TRUE, echo = FALSE}
DAYV2PUB <- read.csv(file = "~/GHI/data/NHTS/Ascii/DAYV2PUB.CSV", na.strings = c("-9","-8", "-7", "-1"))
names(DAYV2PUB) <- c("HOUSEID","PERSONID","FRSTHM","OUTOFTWN","ONTD_P1","ONTD_P2","ONTD_P3","ONTD_P4","ONTD_P5","ONTD_P6","ONTD_P7","ONTD_P8","ONTD_P9","ONTD_P10","ONTD_P11","ONTD_P12","ONTD_P13","ONTD_P14","ONTD_P15","TDCASEID","HH_HISP","HH_RACE","DRIVER","R_SEX","WORKER","DRVRCNT","HHFAMINC","HHSIZE","HHVEHCNT","NUMADLT","FLAG100","LIF_CYC","TRIPPURP","AWAYHOME","CDIVMSAR","CENSUS_D","CENSUS_R","DROP_PRK","DRVR_FLG","EDUC","ENDTIME","HH_ONTD","HHMEMDRV","HHRESP","HHSTATE","HHSTFIPS","INTSTATE","MSACAT","MSASIZE","NONHHCNT","NUMONTRP","PAYTOLL","PRMACT","PROXY","PSGR_FLG","R_AGE","RAIL","STRTTIME","TRACC1","TRACC2","TRACC3","TRACC4","TRACC5","TRACCTM","TRAVDAY","TREGR1","TREGR2","TREGR3","TREGR4","TREGR5","TREGRTM","TRPACCMP","TRPHHACC","TRPHHVEH","TRPTRANS","TRVL_MIN","TRVLCMIN","TRWAITTM","URBAN","URBANSIZE","URBRUR","USEINTST","USEPUBTR","VEHID","WHODROVE","WHYFROM","WHYTO","WHYTRP1S","WRKCOUNT","DWELTIME","WHYTRP90","TDTRPNUM","TDWKND","TDAYDATE","TRPMILES","WTTRDFIN","VMT_MILE","PUBTRANS","HOMEOWN","HOMETYPE","HBHUR","HTRESDN","HTHTNRNT","HTPPOPDN","HTEEMPDN","HBRESDN","HBHTNRNT","HBPPOPDN","GASPRICE","VEHTYPE","HH_CBSA","HHC_MSA")

@@

```{r loadData, eval = FALSE, echo = FALSE}
NHTS.person <- readRDS(file = "./data/NHTS.person.rds")
NHTS.trips <- readRDS(file = "./data/NHTS.trips.rds")
```

@@

```{r summary, fig.width = 20, fig.height = 10, eval = TRUE, echo = FALSE}
summary(with(subset(NHTS.trips, AT == "walk"), TRPMILES))
summary(with(subset(NHTS.trips, AT == "cycle"), TRPMILES))
summary(with(subset(NHTS.trips, AT == "other"), TRPMILES))
summary(with(subset(NHTS.trips, AT == "walk"), TRVLCMIN))
summary(with(subset(NHTS.trips, AT == "cycle"), TRVLCMIN))
summary(with(subset(NHTS.trips, AT == "other"), TRVLCMIN))

@@

```{r subset, eval = TRUE, fig.width = 10, echo = FALSE}
NHTS.person <- subset( NHTS.person, MSA == TRUE )
NHTS <- merge(NHTS.person, NHTS.trips, by = c("HOUSEID","PERSONID"))
barplot(with(NHTS,table(HHS)))
```

@@

```{r split, eval = TRUE, echo = FALSE}
listOfLists <- lapply(split(NHTS, NHTS$HHS),function(x) lapply(split(x, x$SEX), function(x) split(x, x$AGE)))
```

@@

```{r count, eval = TRUE, echo = FALSE}
counts <- lapply(listOfLists, function(x) lapply(x, function(x) lapply(x, countPeople)))
walkLength <- lapply(listOfLists, function(x) lapply(x, function(x) lapply( x, tripTime, mode = "walk")))
cycleLength <- lapply(listOfLists, function(x) lapply(x, function(x) lapply( x, tripTime, mode = "cycle")))
countMatrix <- lapply(counts, reformatData)
walkMatrix <- lapply(walkLength, reformatData)
cycleMatrix <- lapply(cycleLength, reformatData)
meanWalkList <- mapply(walkMatrix,   countMatrix, FUN = function(x,y) x/y*7, SIMPLIFY = FALSE)
meanCycleList <- mapply(cycleMatrix, countMatrix, FUN = function(x,y) x/y*7, SIMPLIFY = FALSE)
RwtList <- lapply( meanWalkList, function(x) ifelse(x == 0, 0.01, x) )
RctList <- lapply( meanCycleList, function(x) ifelse(x == 0, 0.01, x) )
RwtList <- lapply( RwtList, function(x) x/x[3,2] )
RctList <- lapply( RctList, function(x) x/x[3,2] )
muwtList <- lapply( meanWalkList, function(x) x[3,2] )
muctList <- lapply( meanCycleList, function(x) x[3,2] )
```

@@

```{r CDC, eval = TRUE, echo = FALSE}
wonder <- read.csv("./../data/CDCwonder/HHSmetro_Age_sex_2010-14.csv", header=T)
names(wonder) <- c("HHSRegion", "HHSRegionCode", "gender", "genderCode","WONDERage","WONDERageCode","ICD10",  "ICD10Code", "deaths5yr", "population5yr", "crudeRate")

@@

```{r ITHIMCause, eval = TRUE, echo = FALSE}
ITHIMcauseKey <- c("ColonCancer","LungCancer","BreastCancer","Diabetes","Dementia","IHD","InflammatoryHD","InflammatoryHD","HHD","Stroke","AcuteRespInfections","AcuteRespInfections","RespiratoryDisease","RespiratoryDisease","RTIs","RTIs")
names(ITHIMcauseKey) <- unique(wonder$ICD10)
wonder$ITHIMcause <- ITHIMcauseKey[as.character(wonder$ICD10)]
```
```{r recodeAge, eval = TRUE, echo = FALSE}
ITHIMageKey <- c("ageClass1", "ageClass1", "ageClass2", "ageClass2", "ageClass3", "ageClass3", "ageClass3", "ageClass4", "ageClass4", "ageClass4", "ageClass5", "ageClass5", "ageClass5", "ageClass6", "ageClass6", "ageClass7", "ageClass7", "ageClass8", "ageClass8","ageClass8", "ageClass8", "ageClass8")
names(ITHIMageKey) <- unique(wonder$WONDERage)
wonder$ITHIMage <- ITHIMageKey[as.character(wonder$WONDERage)]
```

@@

```{r wonder, eval = TRUE, echo = FALSE}
wonder <- wonder[, c("HHSRegionCode","gender","ITHIMage","ITHIMcause","deaths5yr","population5yr")]
depress <- expand.grid(HHSRegionCode=unique(wonder$HHSRegionCode), gender=unique(wonder$gender),ITHIMage = unique(wonder$ITHIMage), ITHIMcause = "Depression", deaths5yr = "0", population5yr = "0", stringsAsFactors=T)
wonder <- rbind(wonder, depress)
```

@@

```{r suppressed, eval = TRUE, echo = FALSE}
wonder <- within( wonder, {
    newDeaths <- as.numeric(ifelse(deaths5yr == "Suppressed", sample(0:9), as.numeric(as.character(deaths5yr))))
    })
```

@@

```{r summary2, eval = TRUE, echo = FALSE}
wonderSummary <- ddply(wonder, .(HHSRegionCode, gender, ITHIMage, ITHIMcause), summarise,
                       numberOfDeaths5yr = sum(as.numeric(as.character(newDeaths)), na.rm=T)
)
```

@@

```{r singleyear, eval = TRUE, echo = FALSE}
pop14 <- read.csv("./../data/CDCwonder/HHSmetro_Age_sex_2014_AllCause.csv", header=T)

@@

```{r sloppy, eval = TRUE, echo = FALSE}
ITHIMageKey2 <- c("ageClass8", "ageClass1", "ageClass7", "ageClass7", "ageClass6", "ageClass6", "ageClass1", "ageClass5", "ageClass5","ageClass5","ageClass4", "ageClass3", "ageClass4", "ageClass2", "ageClass2", "ageClass3", "ageClass4", "ageClass3","ageClass8")
names(ITHIMageKey2) <- unique(pop14$WONDERage)
pop14$ITHIMage <- ITHIMageKey2[as.character(pop14$WONDERage)]
```

```{r foop, eval = TRUE, echo = FALSE}
pop14 <-  pop14[, c("HHSRegionCode","gender","ITHIMage","deaths2014AllCause","population2014")]
pop14$deaths2014 <- as.numeric(ifelse(pop14$deaths2014 == "Suppressed", sample(0:9), as.numeric(as.character(pop14$deaths2014))))

@@

```{r USGBD, eval = TRUE, echo = FALSE}
USGBD <- read.csv("./../data/CDCwonder/USGBD.csv", header=T)
USGBD$ITHIMage <- as.character(USGBD$ITHIMage)
USGBD$ITHIMcause <- as.character(USGBD$ITHIMcause)

@@

```{r burdenList, eval = TRUE, echo = FALSE}
GBDList <- lapply(split(burdenOut, burdenOut$HHS),function(x) lapply(split(x, x$disease), function(x) split(x, x$sex)))
#saveRDS(GBDList, file = "../data/GBD.rds")
```

@@

```{r newMETs, echo = TRUE, eval = TRUE, echo = FALSE}
newMETsFile <- "~/GHI/data/ATUS/ATUS_to_METS_key.csv"
newMETs <- read.csv(file = newMETsFile, stringsAsFactors = FALSE, colClasses = c("character","character","character","character","numeric","character","character","character"))
dataColumns <- c(1,5)
D <- newMETs[,dataColumns]
metaData <- newMETs[,-dataColumns]
newMETs <- list(Data = D, metaData = metaData)
activityCode <- ifelse(nchar(newMETs$Data$X6.digit.activity.code)==5,paste0("t0",newMETs$Data$X6.digit.activity.code),paste0("t",newMETs$Data$X6.digit.activity.code))

@@

```{r readnewMETs, echo = TRUE, eval = TRUE, echo = FALSE}
cpsFile <- "~/GHI/data/ATUS/atuscps_0311.dat"
cps <- read.table(cpsFile, header=T, skip=0, sep="," )
cps <- within(cps, {
    TUCASEID <- as.character(TUCASEID)
    TULINENO <- as.character(TULINENO)
    GEREG <- as.factor(GEREG)
    })
cps <- subset(cps, TULINENO == "1")

@@

```{r metKey, echo = TRUE, eval = TRUE, echo = FALSE}
allActivities <- colnames(summaryList$Data)
nonTravelActivities <- rownames(newMETs$metaData)[which(newMETs$metaData$Travel.related..Y.N. == "N")]
nonTravelActivities <- nonTravelActivities[which(nonTravelActivities %in% allActivities)]

@@

```{r parameters, eval = TRUE, echo = FALSE}
(parameters <- ddply(D, .(SEX, AGECAT), summarise,
             mean = mean(METs, na.rm=T),
             sd = sd(METs, na.rm=T)
             ))

@@

```{r foobar, eval = TRUE, echo = FALSE}
sumDALY <- function(ITHIM.baseline, ITHIM.scenario){
    return(sum(unlist(compareModels(ITHIM.baseline, ITHIM.scenario)$daly.delta)))
    }
```

```{r ITHIMObjs, eval = TRUE, fig.width = 20, fig.height = 10, echo = FALSE}
compute <- TRUE

@@

```{r meanMatrices, eval = TRUE, fig.width = 20, fig.height = 10, warning = FALSE, message = FALSE, echo = FALSE}

@@

```{r CRA, eval = TRUE, echo = FALSE}

@@

```{r instantiate, eval = TRUE, echo = FALSE}
ITHIM.baseline <- createITHIM(vision = "baseline", region = "SFBayArea")
ITHIM.scenario <- createITHIM(vision = "scenario", region = "SFBayArea")
```

@@

```{r RRs, eval = TRUE, echo = FALSE}
comparativeRisk <- compareModels(ITHIM.baseline,ITHIM.scenario)
```

### Deaths
```{r show, echo = FALSE, eval = TRUE, echo = FALSE}
matrix(unlist(lapply(comparativeRisk$dproj.delta,function(x) lapply(x, function(x) sum(x[-(1:2)]) ))),nrow=5,ncol=2,byrow=TRUE,dimnames=list(names(comparativeRisk$dproj.delta),c("M","F")))
```
### DALY
```{r show2, echo = FALSE, eval = TRUE, echo = FALSE}
matrix(unlist(lapply(comparativeRisk$daly.delta,function(x) lapply(x, function(x) sum(x[-(1:2)]) ))),nrow=5,ncol=2,byrow=TRUE,dimnames=list(names(comparativeRisk$dproj.delta),c("M","F")))
```
### YLD
```{r show3, echo = FALSE, eval = TRUE, echo = FALSE}
matrix(unlist(lapply(comparativeRisk$yld.delta,function(x) lapply(x, function(x) sum(x[-(1:2)]) ))),nrow=5,ncol=2,byrow=TRUE,dimnames=list(names(comparativeRisk$dproj.delta),c("M","F")))
```
### YLL
```{r show4, echo = FALSE, eval = TRUE, echo = FALSE}
matrix(unlist(lapply(comparativeRisk$yll.delta,function(x) lapply(x, function(x) sum(x[-(1:2)]) ))),nrow=5,ncol=2,byrow=TRUE,dimnames=list(names(comparativeRisk$dproj.delta),c("M","F")))
```

@@

```{r instantiate2, eval = TRUE, echo = FALSE}
ITHIM.baseline <- createITHIM(vision = "baseline", region = "national")
ITHIM.scenario <- createITHIM(vision = "scenario", region = "national")
```

```{r RRs2, eval = TRUE, echo = FALSE}
comparativeRisk <- compareModels(ITHIM.baseline,ITHIM.scenario)
```

### Deaths
```{r show222, echo = FALSE, eval = TRUE}
matrix(unlist(lapply(comparativeRisk$dproj.delta,function(x) lapply(x, function(x) sum(x[-(1:2)]) ))),nrow=5,ncol=2,byrow=TRUE,dimnames=list(names(comparativeRisk$dproj.delta),c("M","F")))
```
### DALY
```{r show22, echo = FALSE, eval = TRUE}
matrix(unlist(lapply(comparativeRisk$daly.delta,function(x) lapply(x, function(x) sum(x[-(1:2)]) ))),nrow=5,ncol=2,byrow=TRUE,dimnames=list(names(comparativeRisk$dproj.delta),c("M","F")))
```
### YLD
```{r show23, echo = FALSE, eval = TRUE}
matrix(unlist(lapply(comparativeRisk$yld.delta,function(x) lapply(x, function(x) sum(x[-(1:2)]) ))),nrow=5,ncol=2,byrow=TRUE,dimnames=list(names(comparativeRisk$dproj.delta),c("M","F")))
```
#### YLL
```{r show24, echo = FALSE, eval = TRUE}
matrix(unlist(lapply(comparativeRisk$yll.delta,function(x) lapply(x, function(x) sum(x[-(1:2)]) ))),nrow=5,ncol=2,byrow=TRUE,dimnames=list(names(comparativeRisk$dproj.delta),c("M","F")))
```

@@

```{r example, eval = TRUE, echo = FALSE}
example("ITHIM-package")
```

@@

```{r vision, eval = TRUE, echo = FALSE}
ITHIM.vision <- updateITHIM(ITHIM.scenario, "muwt", 2*ITHIM.scenario$parameters$muwt) # Double the mean walking time
comparativeRisk <- compareModels(ITHIM.baseline,ITHIM.vision)
```
#### YLL
```{r show233, echo = FALSE, eval = TRUE}
matrix(unlist(lapply(comparativeRisk$yll.delta,function(x) lapply(x, function(x) sum(x[-(1:2)]) ))),nrow=5,ncol=2,byrow=TRUE,dimnames=list(names(comparativeRisk$dproj.delta),c("M","F")))
```

@@

```{r params, eval = TRUE, echo = FALSE}
names(ITHIM.baseline$parameters)
ITHIM.baseline$parameters$NonTravelMETs
ITHIM.baseline$parameters$Rwt
```
